Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-05 paper.tex
Reformatted
Added C-alpha macro

\documentclass[12pt]{article}
\bibliographystyle{unsrt}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{url}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

@@

The predictive power of any machine learning model is largely
dependent on the quality and size of the dataset it was trained on. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called abYbank/AbDb\cite{XXXX},
established by Prof. Andrew Martin, was utilised (1924 non-redundant
structures). The root-mean-square deviation (RMSD) value, a measure of
distance between backbone C-$\alpha$ atoms of superimposed crystal structures
and modelled structures, is calculated (see methods). This metric for
model quality was used to classify models.

@@

\subsubsection{Physiochemical Feature Encoding}
In a paper by L. Nanni and A.Lumini a new encoding technique is
presented which was developed for machine learning classifiers. Many
physiochemical properties are calculated and transformed using a
non-linear Fisher transform for dimensionality reduction.  A vector of
length 19 is produced for each amino acid34.  In a paper on designing
a neural network for predicting the packing angle of the light and
heavy variable chain of an antibodies, A.C.R. Martin and
K.R. Abhinandan introduce an encoding method that produces a
four-dimensional physiochemical feature vector25. The amino acid
properties used are 1) the total number of side-chain atoms, 2) the
number of side-chain atoms in the shortest path from C$\alpha$ to the most
distal atom, 3) the Eisenberg consensus
hydrophobicity\cite{Eisenberg1982}, 4) the charge.

@@

\begin{figure}
  \includegraphics[scale=0.6]{angle.eps}
  \caption {Diagram visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e H95 and H102, shown as
    red spheres) of the CDR-H3 (shown in red) are identified. Then, a
    line is drawn between the two C$\alpha$ atoms of these residues (yellow
    dashed line). The distance of the C$\alpha$-atom of each residue in the
    CDR-H3 loop to this line is calculated. The residue which has the
    greatest distance to the line (shown as green sphere) is outputted
    as one-letter amino acid code and used as feature. The distance in
    \AA\ (depicted as yellow arrow) is used as the `protrusion'
    feature. }
  \label{fig:angle}
\end{figure}

>>>>

\documentclass[12pt]{article}
\bibliographystyle{unsrt}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{url}
\newcommand{\ca}{\mbox{C$\alpha$}}

@@

The predictive power of any machine learning model is largely
dependent on the quality and size of the dataset it was trained on. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called abYbank/AbDb\cite{XXXX},
established by Prof. Andrew Martin, was utilised (1924 non-redundant
structures). The root-mean-square deviation (RMSD) value, a measure of
distance between backbone \ca\ atoms of superimposed crystal structures
and modelled structures, is calculated (see methods). This metric for
model quality was used to classify models.

@@

\subsubsection{Physiochemical Feature Encoding}
In a paper by L. Nanni and A.Lumini a new encoding technique is
presented which was developed for machine learning classifiers. Many
physiochemical properties are calculated and transformed using a
non-linear Fisher transform for dimensionality reduction.  A vector of
length 19 is produced for each amino acid34.  In a paper on designing
a neural network for predicting the packing angle of the light and
heavy variable chain of an antibodies, A.C.R. Martin and
K.R. Abhinandan introduce an encoding method that produces a
four-dimensional physiochemical feature vector25. The amino acid
properties used are 1) the total number of side-chain atoms, 2) the
number of side-chain atoms in the shortest path from \ca\ to the most
distal atom, 3) the Eisenberg consensus
hydrophobicity\cite{Eisenberg1982}, 4) the charge.

@@

\begin{figure}
  \includegraphics[scale=0.6]{angle.eps}
  \caption {Diagram visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e H95 and H102, shown as
    red spheres) of the CDR-H3 (shown in red) are identified. Then, a
    line is drawn between the two \ca\ atoms of these residues (yellow
    dashed line). The distance of the \ca-atom of each residue in the
    CDR-H3 loop to this line is calculated. The residue which has the
    greatest distance to the line (shown as green sphere) is outputted
    as one-letter amino acid code and used as feature. The distance in
    \AA\ (depicted as yellow arrow) is used as the `protrusion'
    feature. }
  \label{fig:angle}
\end{figure}
